Against Spurious Understanding: Dovelating Inflated AI Claims \raisebox{-6pt}{\includegraphics[scale=0.09]{dove_1f54a-fe0f.png}}

### 2024-01-31 ECONDAT.tex
Updates from Overleaf

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Submission and Formatting Instructions for ICML 2024}

@@

\subsection{Do neural networks get born with world models?}\label{example-deep-learning}

@@

\subsection{PCA as a yield curve interpreter}\label{example-principal-component-analysis}
%\subsection{Example: Principal Component Analysis}\label{example-principal-component-analysis}

@@

% \begin{figure*}

% \centering{

% \includegraphics[width=1.0\textwidth]{results/figures/dl.png}[tb]

% }

% \caption{\label{fig-dl-results}The left chart shows the actual GDP growth
% and fitted values from the autoencoder model. The right chart shows the
% observed average level and spread of the yield curve (solid) along with
% the predicted values (in-sample) from the linear probe based on the latent embeddings
% (dashed).}

% \end{figure*}%
% %\subsection{Example: Autoencoder}\label{example-deep-learning}
% \subsection{Autoencoders as economic growth predictors}\label{example-deep-learning}

% %So far we have considered simple linear transformations of data. One might argue that the previous examples to not really involve latent embeddings in the way that they are typically thought of in the context of deep learning. Our next example will therefore involve a deep neural network.

% Our goal in this section is to not only predict economic growth from the yield curve, but also extract meaningful features for downstream inference tasks. In particular, we will now use a neural network architecture.
% %, that allows us to recover a compact latent representation of the yield curve.

% \subsubsection{Data}\label{data}

% To estimate economic growth, we will rely on a quarterly
% \href{https://fred.stlouisfed.org/series/GDPC1}{series} of the real gross domestic product (GDP) provided by the Federal Reserve Bank of St.~Louis. The data arrives in terms of levels of real GDP. In order to estimate growth, we transform the data using log differences. Since
% our yield curve data is daily, we aggregate it to the
% quarterly frequency by taking averages of daily yields for each maturity. We also standardize yields since deep learning models tend to perform better with standardized data \cite{gal2019standardization}. Since COVID-19 was a substantial structural break in the time series, we also filter out all observations after 2018.

% \subsubsection{Model}\label{model}

% %Let \(g_t\) denote growth and \(\mathbf{r}_t\) denote the matrix of aggregated Treasury yield rates at time \(t\). Then we are interested in a model for \(g_t\) conditional on \(\mathbf{r}_t\). Let \(\theta\) denote our model parameters then formally we are interested in maximizing the likelihood, \(p_{\theta}(g_t|\mathbf{r}_t)\).

% %To do this, we use a simple autoencoder architecture. With this
% Using a simple autoencoder architecture, we let our model \(g_t\) denote growth and our conditional \(\mathbf{r}_t\) the matrix of aggregated Treasury yield rates at time \(t\). Finally, we let \(\theta\) denote our model parameters. Formally, we are interested in maximizing the likelihood \(p_{\theta}(g_t|\mathbf{r}_t)\).

% %To do this, we use a simple autoencoder architecture: 
% The encoder consists of a single fully connected hidden layer with 32 neurons and a hyperbolic tangent activation function. The bottleneck layer connecting the encoder to the decoder, is a fully connected layer with 6 neurons. The decoder consists of two fully connected layers, each with a hyperbolic tangent activation function: the first layer consists of 32 neurons and the second layer will have the same dimension as the input data. The output layer consists of a single neuron for our output variable, \(g_t\). We train the model over 1,000 epochs to minimize mean squared error loss using the Adam optimizer~\citep{kingma2017adam}.

% The in-sample fit of the model is shown in the left chart of Figure~\ref{fig-dl-results}, which shows actual GDP growth and fitted values from the autoencoder
% model. The model has a large number of free parameters and captures the relationship between economic growth and the yield curve reasonably well, as expected. Since our primary goal is not out-of-sample prediction accuracy but feature extraction for inference, we use all of the available data instead of reserving a hold-out set. As discussed above, we also know that the relationship between economic growth and the yield curve is characterized by two main factors: the level and the spread. Since the model itself is fully characterized by its parameters, we would expect that these two important factors are reflected somewhere in the latent parameter space.

% \subsubsection{Linear Probe}\label{linear-probe}

% While the loss function applies most direct pressure on layers near the final output layer, any information useful for the downstream task first needs to pass through the bottleneck layer \citep{alain2018understanding}. On a per-neuron basis, the pressure to distill useful representation is therefore likely maximized there. Consequently, the bottleneck layer activations seem like a natural place to start looking for compact, meaningful representations of distilled information.  We compute and extract these activations \(A_t\) for all time periods \(t=1,...,T\). Next, we use a linear probe to regress the observed
% yield curve factors on the latent embeddings. Let \(Y_t\) denote the vector containing the two factors of interest in time \(t\): \(y_{t,l}\) and \(y_{t,s}\) for the level and spread, respectively. Formally, we are interested in the following regression model: \(p_{w}(Y_t|A_t)\) where \(w\) denotes the regression
% parameters. We use Ridge regression with \(\lambda\) set to \(0.1\). Using the estimated regression parameters \(\hat{w}\), we then predict the yield curve factors% from the latent embeddings
% : \(\hat{Y}_t=\hat{w}^{\prime}A_t\).

% The in-sample predictions of the probe are shown in the right chart of Figure~\ref{fig-dl-results}. Solid lines show the observed yield curve factors over time, while dashed lines show predicted values. We find that the latent embeddings predict the two yield curve factors reasonably well, in particular the spread. %As in all of the other examples we have presented so, we do not find this result particularly surprising.

% Did the neural network now learn an intrinsic understanding of the economic relationship between growth and the yield curve? To us, that would be too big of statement. Still, the current form of information distillation can be useful, even beyond its intended use for monitoring models.
% %It may serve as a reasonable diagnostic tool to reassure us that the model has picked up on meaningful patterns in the underlying data. But we would certainly not conclude that the model has somehow implicitly developed an intrinsic "understanding" of the economic relationship between growth and the yield curve. While we do not prescribe to this particular view, we still support the idea that this form of information distillation can be useful even beyond its intended use for monitoring models, which we will turn to next.
% %\subsubsection{Potential Use-Case: Inference}\label{ok-but-truly-whats-the-point}
% For example, an interesting idea could be to use the latent embeddings as features in a more traditional and interpretable econometric model. To demonstrate this, let us consider a simple linear regression model for GDP growth. We might be interested in understanding to what degree economic growth in the past is associated with economic growth today. As we might expect, linearly regressing economic growth on lagged growth, as in column (1) of Table \ref{tab-reg}, yields a statistically significant coefficient. But this coefficient suffers from confounding bias, since there are many other confounding variables at play, of which some may be readily observable and measurable, but others may not.

% % A less risky/strong term than confounding bias that we could use here is omitted variable bias. Difference is nuanced: https://stats.stackexchange.com/questions/496328/difference-omitted-variable-bias-and-confounding

% We
% %have, for example, 
% e.g.\ already mentioned the relationship between interest rates and economic growth. To account for that, while keeping our regression model as parsimonious as possible, we could include the level and the spread of the US Treasury yield curve as additional regressors. While this slightly changes the estimated magnitude of the coefficient on lagged growth, the coefficients on the observed level and spread are statistically insignificant (column (2) in Table \ref{tab-reg}). This indicates that these measures may be too crude to capture valuable information about the relationship between yields and economic growth. Because we have included two additional regressors with little to no predictive power, the model fit as measured by the Bayes Information Criterium (BIC) has actually deteriorated.

% Column (3) of Table \ref{tab-reg} shows the effect of instead including one of the latent embeddings that we recovered above in the regression model. In particular, we pick the one latent embedding that we have found to exhibit the most significant effect on the output variable in a separate regression of growth on all latent embeddings. The estimated coefficient on this latent factor is small in magnitude, but statistically significant. The overall model fit, as measured by the BIC as improved and the magnitude of the coefficient on lagged growth, has changed quite a bit. While this is still a very incomplete toy model of economic growth, it appears that the compact latent representation we recovered can  be used in order to mitigate confounding bias.

% \begin{table}
% \caption{Regression output for various models.}\label{tab-reg}%
% \input{results/regression.tex}
% \end{table}

%\subsection{Example: Large Language Model}\label{ex-llm}
\subsection{LLMs for economic sentiment prediction}\label{ex-llm}

@@

Figure~\ref{fig-attack} shows the results of this small test: it shows predicted inflation levels subtracted by \(\mathbb{E}[f(\varepsilon)]\). The median linear probe predictions for sentences about inflation and deflation are indeed substantially higher and lower, respectively than for random noise. Unfortunately, the same is true for sentences about the inflation and deflation in the number of birds, albeit to a somewhat lower degree. This finding holds for both inflation indicators and to a lesser degree also for yields at different maturities, at least qualitatively.

% Next, we propose the following hypothesis test as a minimum viable testing framework to assess if the probe results (may) provide evidence for an actual `understanding' of key economic relationships learned purely from text:

% \begin{proposition}[Parrot
% Test]\protect\hypertarget{prp-line}{}\label{prp-line}

% ~

% \begin{itemize}
% \setlength\itemsep{1px}
% \item
%   \emph{H0} (Null \emph{WEAK}): The probe never predicts values that are statistically significantly different from \(\mathbb{E}[f(\varepsilon)]\).
% \item
%   \emph{H1} (Stochastic Parrots \emph{RISKIER}): The probe predicts values that are statistically significantly different from \(\mathbb{E}[f(\varepsilon)]\) for sentences related to the outcome of interest \emph{and} those that are independent (i.e. sentences in all categories).
% \item
%   \emph{H2} (More than Mere Stochastic Parrots \emph{RISKIEST}): The probe predicts values that are statistically significantly different from \(\mathbb{E} [f(\varepsilon)]\) for sentences in that are related to the outcome variable (IP and DP), but not for sentences that are independent of the outcome (IB and DB).
% \end{itemize}
% \end{proposition}

% To be clear, if in such a test we did find substantial evidence in favour of rejecting both \emph{HO} and \emph{H1}, this would not automatically imply that \emph{H2} is true. But to even continue investigating, if based on having learned meaningful representation the underlying LLM is more than just a parrot, it should be able to pass this simple test.

% In this particular case, Figure~\ref{fig-attack} demonstrates that we find some evidence to reject \emph{H0} but not \emph{H1} for \emph{FOMC-RoBERTa}. The median linear probe predictions for sentences about inflation and deflation are indeed substantially higher and lower, respectively than for random noise. Unfortunately, the same is true for sentences about the inflation and deflation in the number of birds, albeit to a somewhat lower degree. This finding holds for both inflation indicators and to a lesser degree also for yields at different maturities, at least qualitatively.

% We should note that the number of sentences in each category is very small here (10), so the results in Figure~\ref{fig-attack} cannot be used to establish statistical significance. That being said, even a handful of convincing counter-examples should be enough for us to seriously question the claim, that results from linear probes provide evidence in favor of real `understanding'. In fact, even a handful of sentences for which any human annotator would easily arrive at the conclusion of independence, a prediction by the probe in either direction casts doubt.

\section{On human proneness to over-interpretation}
Linear probes and related tools from mechanistic interpretability were proposed in the context of monitoring models and diagnosing potential problems \citep{alain2018understanding}. Favorable outcomes from probes merely indicate that the model ``has learned information relevant for the property [of interest]" \citep{belinkov2021probing}. Our examples demonstrate, that this is achievable even for small models, while these models have certainly not developed an intrinsic ``understanding'' of the world. Thus, to draw conclusions about the emerging capabilities of AI models, we argue that more conservative and rigorous tests are needed.

@@

\subsection{Spurious relationships}
In statistics, misleading patterns are often referred to as spurious relationships: associations, often quantitatively assessed, between two or more variables that are not causally related to each other. Although the formal definition of spuriousness varies somewhat \cite{haig2003spurious}, it distinctly implies that the observation of correlations does not necessarily imply causation. Quantitative data often show non-causal associations (as humorously demonstrated on the \href{http://www.tylervigen.com/spurious-correlations}{Spurious Correlations} website), and as adept as humans are at recognizing patterns, we typically have a much harder time discerning spurious relationships from causal ones. A major contributor is that humans struggle to tell the difference between random and non-random sequences \cite{falk1997making}, and to generate sequences that appear random \cite{ladouceur1996erroneous}. A common issue is a lack of expectation that randomness that hints towards a causal relationship, such as correlations, will still appear at random. This leads even those trained in statistics and probability to perceive illusory correlations, correlations of inflated magnitude (see \citet{nickerson1998confirmation}), or causal relationships in data that is randomly generated \cite{zgraggen2018investigating}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix
\onecolumn
\section{You \emph{can} have an appendix here.}

You can have as much text here as you want. The main body must be at most $8$ pages long.
For the final version, one more page can be added.
If you want, you can use an appendix like this one.

The $\mathtt{\backslash onecolumn}$ command above can be kept in place if you prefer a one-column appendix, or can be removed if you prefer a two-column appendix.  Apart from this possible change, the style (font size, spacing, margins, page numbering, etc.) should be kept the same as the main body.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Against Spurious Understanding: Dovelating Inflated AI Claims}

@@

\subsection{A Neural Networks Born with World Models?}\label{example-deep-learning}

@@

\subsection{PCA as a Yield Curve Interpreter}\label{example-principal-component-analysis}
%\subsection{Example: Principal Component Analysis}\label{example-principal-component-analysis}

@@

\includegraphics[width=1.0\textwidth]{results/figures/dl.png}[tb]

\caption{\label{fig-dl-results}The left chart shows the actual GDP growth
and fitted values from the autoencoder model. The right chart shows the
observed average level and spread of the yield curve (solid) along with
the predicted values (in-sample) from the linear probe based on the latent embeddings
(dashed).}

\end{figure*}%
%\subsection{Example: Autoencoder}\label{example-deep-learning}
\subsection{Autoencoders as Economic Growth Predictors}\label{example-deep-learning}

%So far we have considered simple linear transformations of data. One might argue that the previous examples to not really involve latent embeddings in the way that they are typically thought of in the context of deep learning. Our next example will therefore involve a deep neural network.

Our goal in this section is to not only predict economic growth from the yield curve, but also extract meaningful features for downstream inference tasks. In particular, we will now use a neural network architecture.
%, that allows us to recover a compact latent representation of the yield curve.

\subsubsection{Data}\label{data}

To estimate economic growth, we will rely on a quarterly
\href{https://fred.stlouisfed.org/series/GDPC1}{series} of the real gross domestic product (GDP) provided by the Federal Reserve Bank of St.~Louis. The data arrives in terms of levels of real GDP. In order to estimate growth, we transform the data using log differences. Since
our yield curve data is daily, we aggregate it to the
quarterly frequency by taking averages of daily yields for each maturity. We also standardize yields since deep learning models tend to perform better with standardized data \cite{gal2019standardization}. Since COVID-19 was a substantial structural break in the time series, we also filter out all observations after 2018.

\subsubsection{Model}\label{model}

%Let \(g_t\) denote growth and \(\mathbf{r}_t\) denote the matrix of aggregated Treasury yield rates at time \(t\). Then we are interested in a model for \(g_t\) conditional on \(\mathbf{r}_t\). Let \(\theta\) denote our model parameters then formally we are interested in maximizing the likelihood, \(p_{\theta}(g_t|\mathbf{r}_t)\).

%To do this, we use a simple autoencoder architecture. With this
Using a simple autoencoder architecture, we let our model \(g_t\) denote growth and our conditional \(\mathbf{r}_t\) the matrix of aggregated Treasury yield rates at time \(t\). Finally, we let \(\theta\) denote our model parameters. Formally, we are interested in maximizing the likelihood \(p_{\theta}(g_t|\mathbf{r}_t)\).

%To do this, we use a simple autoencoder architecture: 
The encoder consists of a single fully connected hidden layer with 32 neurons and a hyperbolic tangent activation function. The bottleneck layer connecting the encoder to the decoder, is a fully connected layer with 6 neurons. The decoder consists of two fully connected layers, each with a hyperbolic tangent activation function: the first layer consists of 32 neurons and the second layer will have the same dimension as the input data. The output layer consists of a single neuron for our output variable, \(g_t\). We train the model over 1,000 epochs to minimize mean squared error loss using the Adam optimizer~\citep{kingma2017adam}.

The in-sample fit of the model is shown in the left chart of Figure~\ref{fig-dl-results}, which shows actual GDP growth and fitted values from the autoencoder
model. The model has a large number of free parameters and captures the relationship between economic growth and the yield curve reasonably well, as expected. Since our primary goal is not out-of-sample prediction accuracy but feature extraction for inference, we use all of the available data instead of reserving a hold-out set. As discussed above, we also know that the relationship between economic growth and the yield curve is characterized by two main factors: the level and the spread. Since the model itself is fully characterized by its parameters, we would expect that these two important factors are reflected somewhere in the latent parameter space.

\subsubsection{Linear Probe}\label{linear-probe}

While the loss function applies most direct pressure on layers near the final output layer, any information useful for the downstream task first needs to pass through the bottleneck layer \citep{alain2018understanding}. On a per-neuron basis, the pressure to distill useful representation is therefore likely maximized there. Consequently, the bottleneck layer activations seem like a natural place to start looking for compact, meaningful representations of distilled information.  We compute and extract these activations \(A_t\) for all time periods \(t=1,...,T\). Next, we use a linear probe to regress the observed
yield curve factors on the latent embeddings. Let \(Y_t\) denote the vector containing the two factors of interest in time \(t\): \(y_{t,l}\) and \(y_{t,s}\) for the level and spread, respectively. Formally, we are interested in the following regression model: \(p_{w}(Y_t|A_t)\) where \(w\) denotes the regression
parameters. We use Ridge regression with \(\lambda\) set to \(0.1\). Using the estimated regression parameters \(\hat{w}\), we then predict the yield curve factors% from the latent embeddings
: \(\hat{Y}_t=\hat{w}^{\prime}A_t\).

The in-sample predictions of the probe are shown in the right chart of Figure~\ref{fig-dl-results}. Solid lines show the observed yield curve factors over time, while dashed lines show predicted values. We find that the latent embeddings predict the two yield curve factors reasonably well, in particular the spread. %As in all of the other examples we have presented so, we do not find this result particularly surprising.

Did the neural network now learn an intrinsic understanding of the economic relationship between growth and the yield curve? To us, that would be too big of statement. Still, the current form of information distillation can be useful, even beyond its intended use for monitoring models.

For example, an interesting idea could be to use the latent embeddings as features in a more traditional and interpretable econometric model. To demonstrate this, let us consider a simple linear regression model for GDP growth. We might be interested in understanding to what degree economic growth in the past is associated with economic growth today. As we might expect, linearly regressing economic growth on lagged growth, as in column (1) of Table \ref{tab-reg}, yields a statistically significant coefficient. But this coefficient suffers from confounding bias, since there are many other confounding variables at play, of which some may be readily observable and measurable, but others may not.

We e.g.\ already mentioned the relationship between interest rates and economic growth. To account for that, while keeping our regression model as parsimonious as possible, we could include the level and the spread of the US Treasury yield curve as additional regressors. While this slightly changes the estimated magnitude of the coefficient on lagged growth, the coefficients on the observed level and spread are statistically insignificant (column (2) in Table \ref{tab-reg}). This indicates that these measures may be too crude to capture valuable information about the relationship between yields and economic growth. Because we have included two additional regressors with little to no predictive power, the model fit as measured by the Bayes Information Criterium (BIC) has actually deteriorated.

Column (3) of Table \ref{tab-reg} shows the effect of instead including one of the latent embeddings that we recovered above in the regression model. In particular, we pick the one latent embedding that we have found to exhibit the most significant effect on the output variable in a separate regression of growth on all latent embeddings. The estimated coefficient on this latent factor is small in magnitude, but statistically significant. The overall model fit, as measured by the BIC as improved and the magnitude of the coefficient on lagged growth, has changed quite a bit. While this is still a very incomplete toy model of economic growth, it appears that the compact latent representation we recovered can  be used in order to mitigate confounding bias.

\begin{table}
\caption{Regression output for various models.}\label{tab-reg}%
\input{results/regression.tex}
\end{table}

%\subsection{Example: Large Language Model}\label{ex-llm}
\subsection{LLMs for Economic Sentiment Prediction}\label{ex-llm}

@@

Next, we propose the following hypothesis test as a minimum viable testing framework to assess if the probe results (may) provide evidence for an actual `understanding' of key economic relationships learned purely from text:

\begin{proposition}[Parrot
Test]\protect\hypertarget{prp-line}{}\label{prp-line}

~

\begin{itemize}
\setlength\itemsep{1px}
\item
  \emph{H0} (Null): The probe never predicts values that are statistically significantly different from \(\mathbb{E}[f(\varepsilon)]\).
\item
  \emph{H1} (Stochastic Parrots): The probe predicts values that are statistically significantly different from \(\mathbb{E}[f(\varepsilon)]\) for sentences related to the outcome of interest \emph{and} those that are independent (i.e. sentences in all categories).
\item
  \emph{H2} (More than Mere Stochastic Parrots): The probe predicts values that are statistically significantly different from \(\mathbb{E} [f(\varepsilon)]\) for sentences in that are related to the outcome variable (IP and DP), but not for sentences that are independent of the outcome (IB and DB).
\end{itemize}
\end{proposition}

To be clear, if in such a test we did find substantial evidence in favour of rejecting both \emph{HO} and \emph{H1}, this would not automatically imply that \emph{H2} is true. But to even continue investigating, if based on having learned meaningful representation the underlying LLM is more than just a parrot, it should be able to pass this simple test.

In this particular case, Figure~\ref{fig-attack} demonstrates that we find some evidence to reject \emph{H0} but not \emph{H1} for \emph{FOMC-RoBERTa}. The median linear probe predictions for sentences about inflation and deflation are indeed substantially higher and lower, respectively than for random noise. Unfortunately, the same is true for sentences about the inflation and deflation in the number of birds, albeit to a somewhat lower degree. This finding holds for both inflation indicators and to a lesser degree also for yields at different maturities, at least qualitatively.

We should note that the number of sentences in each category is very small here (10), so the results in Figure~\ref{fig-attack} cannot be used to establish statistical significance. That being said, even a handful of convincing counter-examples should be enough for us to seriously question the claim, that results from linear probes provide evidence in favor of real `understanding'. In fact, even a handful of sentences for which any human annotator would easily arrive at the conclusion of independence, a prediction by the probe in either direction casts doubt.

\section{On Human Proneness to Over-Interpretation}
Linear probes and related tools from mechanistic interpretability were proposed in the context of monitoring models and diagnosing potential problems \citep{alain2018understanding}. Favorable outcomes from probes merely indicate that the model ``has learned information relevant for the property [of interest]" \citep{belinkov2021probing}. Our examples demonstrate, that this is achievable even for small models, while these models have certainly not developed an intrinsic ``understanding'' of the world. Thus, to draw conclusions about the emerging capabilities of AI models, we argue that more conservative and rigorous tests are needed.

@@

\subsection{Spurious Relationships}
In statistics, misleading patterns are often referred to as spurious relationships: associations, often quantitatively assessed, between two or more variables that are not causally related to each other. Although the formal definition of spuriousness varies somewhat \cite{haig2003spurious}, it distinctly implies that the observation of correlations does not necessarily imply causation. Quantitative data often show non-causal associations (as humorously demonstrated on the \href{http://www.tylervigen.com/spurious-correlations}{Spurious Correlations} website), and as adept as humans are at recognizing patterns, we typically have a much harder time discerning spurious relationships from causal ones. A major contributor is that humans struggle to tell the difference between random and non-random sequences \cite{falk1997making}, and to generate sequences that appear random \cite{ladouceur1996erroneous}. A common issue is a lack of expectation that randomness that hints towards a causal relationship, such as correlations, will still appear at random. This leads even those trained in statistics and probability to perceive illusory correlations, correlations of inflated magnitude (see \citet{nickerson1998confirmation}), or causal relationships in data that is randomly generated \cite{zgraggen2018investigating}.

@@
